Towards a Complete Model of Mutation Selection Balance in Humans

### 2023-03-21 manuscript/manuscript.tex
phrasing
minor ms changes
ms updates

whole-genome variance models.

@@

@@

\begin{align}
  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
\end{align}

@@

Using $Z_\text{BGS}$, $V_m$, and $V$ under the background selection model, and
approximation $N_{e,\infty}^\text{SC98}$ (equation XXX), find an equation
reminiscent of the classic background selection with recombination results. We
find that by using the full equation XXX $N_{e,\infty}$ rather than the
approximation of Santiago and Caballero, the reduction due to background
selection is

\begin{align}
  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right)
\end{align}

which matches equation 8 of \textcite{Hudson1995-xc} except for the $Ls$ term
and equation 10 of \textcite{Nordborg1996-nq} except for the $2s$ term.

@@

@@

While the selection dynamics are set by the asymptotic $Q_\infty^2$ that
determines the equilibrium $\widetilde{N}_e$ and $\widetilde{R}$,
heterozygosity is determined by the sequence $\widetilde{N}_{e,t}$, where each
term is

XXX

>>>>

whole-genome variance models.
These are flux models.

@@

Evolutionary processes describe how genetic variation is introduced and
ultimately converted to substitutions or lost from the population. Our model 
makes predictions about the eventual outcome of weakly deleterious mutation 
that seem to roughly match observations.

@@

\begin{align}
  \label{eq:Ne_inf}
  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
\end{align}

@@

Now, we can use these values for $Z_\text{BGS}$, $V_m$, and $V$ in equation
\eqref{eq:Ne_inf}. Through this alternate quantitative genetics derivation, we
arrive at the same equation as classic BGS theory,

\begin{align}
  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right).
\end{align}

This matches equation 8 of \textcite{Hudson1995-xc} except for the $Ms$ term
and equation 10 of \textcite{Nordborg1996-nq} except for the $2s$ term.

@@

\subsection*{Misspecification of the Polygenic Selection Model}

Thus far, we have developed the general model of equation \eqref{XXX} by
assuming that the only source of $V_h$ is from the deleterious
mutation-selection-draft-drift process. However, other polygenic evolutionary
processes can also create, alter, and eliminate fitness variation. By assuming
that the only fitness variation originates from deleterious
mutation-selection-drift-draft processes, we introduce some model
misspecification. There are two main types of misspecification under the
selection model. First, non-polygenic positive selection processes, such as
recurrent sweeps, may create small-scale distortions in local genealogies.
These have been found to only impact diversity on small spatial scales
\parencite{Murphy2022-sj} in humans, and explain little of the total variation
in megabase-scale patterns of diversity. Second, there is model
misspecification due to our model assuming that the only polygenic selection
process happening is the flux of new deleterious mutations in, and their
eventual loss or fixation. Other polygenic selection processes could be
occurring, e.g. traits could be responding to directional selection. The
Santiago and Caballero model our work extends is very general, and can account
for multiple types of polygenic selection. The key input parameters are the
equilibrium heritable fitness variation $V_h$, and the variation introduced by
mutation each generation $V_m$.

Let us assume for now that,

\begin{align}
  \label{eq:Ne_inf}
  N_{e,\infty} \approx N \exp\left(-\frac{V_h + \varepsilon}{(1-Z) (2-(2-M) Z)} \right).
\end{align}

@@

While the selection dynamics are set by the asymptotic $Q_\infty^2$ that
determines the equilibrium $\widetilde{N}_e$ and $\widetilde{R}$,
heterozygosity is determined by the sequence $\widetilde{N}_{e,t}$, where each
term is given by \eqref{eq:net_integral}. We experimented with calculating this
full sum in our methods, but it seemed to make little difference in our
application. Additionally, it was too costly to calculate computationally
for genome-wide calculations of the B' map.
